---
title: "p8105_hw2_yc4617"
output: github_document
author: Yutong Chen
date: "2024-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 0:
## Read csv file
```{r}
data = read.csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
head(data)
```
```{r}
library(dplyr)
library(tidyverse)
```

## Convert Entry from charater to logical variable
```{r}
data <- data %>%
  mutate(Entry = case_match(Entry, 
                            "YES" ~ TRUE, 
                            "NO" ~ FALSE))
data <- select(data,Line:Route11,Entry,Vending,Entrance.Type,ADA)
head(data)
```
## Describe the dataset
The data set contains these variables: `r paste(names(data), collapse = ", ")`. Until now, I read the data, then convert Entry variable from charater to logical variable, select several important variables from it, and show the variables' names. The resulting data set has `r dim(data)[1]` rows and `r dim(data)[2]` columns. These data are not tidy because there are some NA in column Route 8 - Route 11.

# Problem 1:
## Distinct station
```{r}
number_distinct_station = data %>%
  distinct(Line, Station.Name)%>%
  nrow()
number_distinct_station
```
There are `r number_distinct_station` distinct stations.

## ADA compliant
```{r}
number_ADA_compliant = data %>%
  filter(ADA == TRUE)%>%
  distinct(Line, Station.Name)%>%
  nrow()
number_ADA_compliant
```
There are `r number_ADA_compliant` stations that are ADA compliant.

## Proportion of entrances
```{r}
proportion_entrances = data %>%
  filter(Vending == "NO")%>%
  summarise(rate = mean(Entry == "YES")) %>%
  pull(rate)
```
There are `r proportion_entrances` proportion of station entrances / exits without vending allow entrance.

## ADA in Route A
```{r}
library(tidyverse)

data_reformat = data %>%
  mutate(across(starts_with("Route"), as.character))%>%
  pivot_longer(cols = starts_with("Route"), 
               names_to = "Route_Number", 
               values_to = "Route")
data_reformat
number_A = data_reformat %>%
  filter(Route == "A") %>%
  distinct(Line, Station.Name) %>%
  nrow()
number_A_ADA = data_reformat %>%
  filter(Route == "A", ADA == TRUE) %>%
  distinct(Line, Station.Name) %>%
  nrow()
```

There are `r number_A` distinct stations serve the A train. Of the stations that serve the A train, there are `r number_A_ADA` ADA compliant.

# Problem 2:
## Preprocess files
```{r}
library(readxl)

Mr_Trash_wheel = read_excel("202409 Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel") %>%
  select(where(~!all(is.na(.)))) %>%
  rename(Weight_tons = "Weight (tons)",
         Volume_cubic_yards = "Volume (cubic yards)",
         Plastic_Bottles = "Plastic Bottles",
         Cigarette_Butts = "Cigarette Butts",
         Glass_Bottles = "Glass Bottles",
         Plastic_Bags = "Plastic Bags",
         Sports_Balls = "Sports Balls",
         Homes_Powered = "Homes Powered*") %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Sports_Balls = as.integer(round(Sports_Balls))) %>%
  mutate(Trash_wheel = "Mr.Trash Wheel") %>%
  mutate(Year = as.character(Year))

Professor_Trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx",sheet = "Professor Trash Wheel") %>%
  select(where(~!all(is.na(.)))) %>%
  rename(Weight_tons = "Weight (tons)",
         Volume_cubic_yards = "Volume (cubic yards)",
         Plastic_Bottles = "Plastic Bottles",
         Cigarette_Butts = "Cigarette Butts",
         Glass_Bottles = "Glass Bottles",
         Plastic_Bags = "Plastic Bags",
         Homes_Powered = "Homes Powered*") %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Trash_wheel = "Professor Trash Wheel")%>%
  mutate(Year = as.character(Year))

Gwynnda_Trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx",sheet = "Gwynnda Trash Wheel") %>%
  select(where(~!all(is.na(.)))) %>%
  rename(Weight_tons = "Weight (tons)",
         Volume_cubic_yards = "Volume (cubic yards)",
         Plastic_Bottles = "Plastic Bottles",
         Cigarette_Butts = "Cigarette Butts",
         Plastic_Bags = "Plastic Bags",
         Homes_Powered = "Homes Powered*") %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Trash_wheel = "Gwynnda Trash Wheel")%>%
  mutate(Year = as.character(Year))
```

## Combine files
```{r}
Combined_Trash_wheel = bind_rows(Mr_Trash_wheel,Professor_Trash_wheel,Gwynnda_Trash_wheel)

Weight_Professor = Combined_Trash_wheel %>%
  filter(Trash_wheel == "Professor Trash Wheel") %>%
  summarise(result = sum(Weight_tons)) %>%
  pull(result)

Number_Cigarette = Combined_Trash_wheel %>%
  filter(Trash_wheel == "Gwynnda Trash Wheel", Month == "June", Year == "2022") %>%
  summarise(result = sum(Cigarette_Butts)) %>%
  pull(result)
Number_Cigarette
```
The total weight of trash collected by Professor Trash Wheel is `r Weight_Professor`. And the total number of cigarette butts collected by Gwynnda in June of 2022 is `r Number_Cigarette`.

# Problem 3:
## Clean files
```{r}
Bakers = read_csv("bakers.csv")
Bakers = janitor::clean_names(Bakers)
Bakes = read_csv("bakes.csv")
Bakes = janitor::clean_names(Bakes)
Results = read_csv("results.csv", skip = 2)
```
```{r}
library(stringr)
Bakes = Bakes %>%
  mutate(baker = str_replace_all(baker, '"', ''))

Bakers <- Bakers %>%
  mutate(baker_name = word(baker_name, 1))

missing_bakers <- anti_join(Bakes, Bakers, by = c('series', 'baker' = 'baker_name'))
print(missing_bakers)
```

```{r}
Combined = Bakes %>%
  left_join(Bakers,by = c('series', 'baker' = "baker_name"))
Combined = Combined %>%
  left_join(Results, by = c('series', 'episode', 'baker'))

Combined <- Combined %>%
  mutate(across(c("technical"), ~replace_na(., -1))) %>%
  mutate(across(c("result"), ~replace_na(., "UNKNOWN")))

write_csv(Combined, "Combined_bakers_dataset.csv")
```
Cleaning steps: When I import these data sets, firstly clean their variables names by using janitor::clean_names. For results.csv, I skip two first unnecessary rows. And then I remove the signal "" in baker names in bakes.csv and extract first name of baker names in bakers.csv to make them the same. After that, I check if there is any missing bakers in two files. After merging these three files, I handle the NA in technical and result columns by replacing NA with -1 and "UNKNOWN" based on their data type. 

## Winners from season 5 to 10 
```{r}
Combined
Winner_bakers <- Combined %>%
  filter(series >= 5 & series <= 10) %>%
  filter(result %in% c("STAR BAKER", "WINNER")) %>%
  select(series, episode, baker, result)
```
Comment: There is no predictable overall winners because each winner only appear once and they have the same possibility to be a overall winner. In each series, intuitively, winner of the series should be the one who frequently get star baker. In fact, although Richard got star baker in series 5 for 5 times, winner of series 5 is Nancy. That's surprising.

## Average of season 1 and 5
```{r}
Viewers = read_csv("viewers.csv")

Season_1_avg <- Viewers %>%
  pull(`Series 1`) %>% 
  mean(na.rm = TRUE)

Season_5_avg <- Viewers %>%
  pull(`Series 5`) %>% 
  mean(na.rm = TRUE)

```
The average viewership in Season 1 is `r Season_1_avg`. In Season 5, the average viewership is `r Season_5_avg`.



